R Notebook

### 2021-11-02 FEM_R/Wealth_Health_Socioeconomics.Rmd
Added visualisations for the first replacement scenario (risk behaviours switched with quint3) (some visualisations are broken). Also generated a new replacement population, where risk behaviours, disability status, and health status are switched from quintile 4 onto quintile 1

@@

rm(imp, test, test.1, test.3)
```

@@

```{r}
## FINALLY SAVE THE DATA
write_dta(repl.final2, 'input_data/ELSA_repl_modified_quint1.dta', version = 12)
write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)
```

@@

## Data
```{r}

@@

## Survival Curve
```{r}

# Function to compare survival curves between quintiles

## THIS IS NOT USEFUL!!!
# Because we calculate survival using the m_endpop_* variable (which is a 
# population level variable) it will always give the same value for every
# subgroup.
# Need to talk to Bryan about a way to analyse this within subgroup.
survivalCurveSubgroup <- function(cohort) {

@@

## Prevalences
```{r}

@@

  # Select just the things we want
  #prev.all <- cohort %>% select(Age, p_anydisease_wq1, p_anydisease_wq2, p_anydisease_wq3, p_anydisease_wq4, p_anydisease_wq5)
  prev.all <- cohort %>% select(Age, all_of(varnames))

  # # pivot the data to long format for ggplot
  # survivals <- pivot_longer(data = prev.all,
  #                           cols = p_anydisease_wq1:p_anydisease_wq5,
  #                           names_to = "Quintile",
  #                           values_to = "Disease_Prev")

@@

>>>>

#### Risk Behaviours only

@@

rm(imp, test, test.1, test.3)

@@

```{r}
## FINALLY SAVE THE DATA
write_dta(repl.final2, 'input_data/ELSA_repl_risk_quint3.dta', version = 12)
write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)

rm(repl.final2, repl2, hotdeck.vars, demog.vars, risk.vars, id.var,
   repl.final, test2)
```

#### Risk Behaviours, ADLs, and Health Status

```{r}
# Set up lists of id vars, demographic vars (for imputation) and risk vars to be imputed
id.var <- 'idauniq'
demog.vars <- c('wealth_quint', 'age', 'male')
change.vars <- c('smokev', 'l2smokev', 'smoken', 'l2smoken', 'heavy_smoker',
               'l2heavy_smoker', 
               'drink', 'l2drink', 'problem_drinker', 'l2problem_drinker', 
               'logbmi', 'l2logbmi', 
               'exstat', 'l2exstat', 
               'adlstat', 'l2adlstat', 'iadlstat', 'l2iadlstat',
               'cancre', 'l2cancre', 'diabe', 'l2diabe', 'hearte', 'l2hearte',
               'lunge', 'l2lunge', 'hibpe', 'l2hibpe', 'stroke', 'l2stroke',
               'asthmae', 'l2asthmae')

test <- repl %>% filter(year == 2012, wealth_quint %in% c(1, 4)) %>%
                  select(all_of(id.var), all_of(demog.vars), all_of(change.vars))

# separate test df by quintile so we can remove data for quint1
test.1 <- test %>% filter(wealth_quint == 1)
test.4 <- test %>% filter(wealth_quint == 4)
test.1[change.vars] <- NA
test <- rbind(test.1, test.4)

# Run mi & get a complete version of the imputed dataset
imp <- mice(data=test, m=5, maxit=10, printFlag = FALSE)
imputed <- complete(imp)

rm(imp, test, test.1, test.4)

```{r}
### ACCOUNTING
## The imputation doesn't work perfectly, and a lot of the lagged variables are not imputed in that step as they are correlated with their current

# copy so we don't mess with original
accounted <- imputed

## CHRONIC CONDITIONS
# Handle chronic conditions first as they are the easiest to impute
absorbing_vars <- c('cancre', 'diabe', 'hearte',
                     'lunge', 'hibpe', 'stroke',
                     'asthmae')

## If var is absorbing, then 0 in var.current must also be 0 in var.lagged
for (var in absorbing_vars) {
  accounted[[paste0('l2', var)]][accounted[[var]] == 0] <- 0
  accounted[[paste0('l2', var)]][accounted[[var]] == 1] <- 1
}

## specific issue with stroke
# Stroke didn't impute, but looking at data for the quantile we swapped with
# not a single case of stroke, so going to set all to 0
accounted$stroke <- 0
accounted$l2stroke <- 0

## ADLs/IADLs
# if adlstat or iadlstat == 1, then lag variable should also be 1
# i.e. if no disabilitys in current, then previous also no disabilities
accounted$l2adlstat[accounted$adlstat == 1] <- 1
accounted$l2iadlstat[accounted$iadlstat == 1] <- 1

## RISK BEHAVIOURS
# if smoken == 1 then most likely l2smoken == 1 also
accounted$l2smoken[is.na(accounted$l2smoken)] <- accounted$smoken[is.na(accounted$l2smoken)]

# if l2smokev == 1 then smokev must also == 1
accounted$smokev[is.na(accounted$smokev)] <- accounted$l2smokev[is.na(accounted$smokev)]
# also, if l2smoken == 1 then l2smokev must also equal 1
accounted$l2smokev[accounted$l2smoken == 1] <- 1
# and if l2smokev == 1 then smokev == 1
accounted$smokev[accounted$l2smokev == 1] <- 1
# same for smoken
accounted$l2smoken[accounted$smoken == 1] <- 1
# going to assume that someone who is a heavy smoker now was a heavy smoker in lag as well
accounted$l2heavy_smoker[is.na(accounted$l2heavy_smoker)] <- accounted$heavy_smoker[is.na(accounted$l2heavy_smoker)]

# if person is a problem drinker then they must also be drink == 1
accounted$drink[accounted$l2problem_drinker == 1] <- 1
# probably unlikely that someone would change drink status, so set lag == current
accounted$l2drink[is.na(accounted$l2drink)] <- accounted$drink[is.na(accounted$l2drink)]
# no variability in problem_drinker -> l2problem_drinker in median wealth group,
# so assume same here
accounted$l2problem_drinker[is.na(accounted$l2problem_drinker)] <- accounted$problem_drinker
```

# copy this so we don't mess with the original repl
repl.quint1 <- repl %>% filter(wealth_quint == 1) %>%
                          select(-all_of(change.vars), -all_of(demog.vars))

# get just the quint 1 people
quint1.replaced <- accounted %>% filter(wealth_quint == 1)

# merge the modified repl with the newly replaced quint1 people
repl.replaced <- merge(repl.quint1, quint1.replaced, by='idauniq', all.x = TRUE)

rm(repl.quint1, quint1.replaced, repl.replaced, repl2, accounted, imputed)
```

## IMPUTING DATA TO RUN THROUGH MODEL

repl.final2 <- hotdeck(repl.final2, variable = hotdeck.vars)
repl2 <- hotdeck(repl2, variable = hotdeck.vars)

rm(hotdeck.vars, var)
```

```{r}
## FINALLY SAVE THE DATA
write_dta(repl.final2, 'input_data/ELSA_repl_rdh_quint4.dta', version = 12)
write_dta(repl2, 'input_data/ELSA_repl_standard.dta', version = 12)

rm(repl.final, repl.final2, repl2, absorbing_vars, change.vars, demog.vars, id.var)
```

@@

## Baseline

### Data
```{r}

@@

### Survival Curve
```{r}

# Function to compare survival curves between quintiles
survivalCurveSubgroup <- function(cohort) {

@@

### Prevalences
```{r}

@@

@@

## Replacement Scenarios

```{r}
cohort <- read_dta("output/WEALTH/ELSA_cohort/ELSA_cohort_summary.dta")
replaced1 <- read_dta("output/WEALTH/ELSA_risk_quint3/ELSA_risk_quint3_summary.dta")
replaced2 <- read_dta("output/WEALTH/ELSA_rdh_quint4/ELSA_rdh_quint4_summary.dta")
```

### Survival Curve
```{r "quint3"}

survivalCurveReplaced <- function(cohort, replaced) {

  # Age and survival for each quintile
  cohort$Age <- (cohort$year - 2012) + 50
  cohort$q1 <- cohort$m_endpop_wq1 / cohort$m_endpop_wq1[1]
  cohort$q3 <- cohort$m_endpop_wq3 / cohort$m_endpop_wq3[1]
  replaced$Age <- (replaced$year - 2012) + 50
  replaced$q1 <- replaced$m_endpop_wq1 / replaced$m_endpop_wq1[1]

  # Cut dataframe down to just the age and survivals
  #survivals <- cohort %>% select(Age, q1, q2, q3, q4, q5)
  survivals <- data.frame(cohort$Age, cohort$q1, cohort$q3, replaced$q1)

  survivals <- survivals %>%
                  dplyr::rename('Age' = cohort.Age, 
                                'CohortQ1' = cohort.q1, 
                                'CohortQ3' = cohort.q3, 
                                'ReplacedQ1' = replaced.q1) %>%
                  pivot_longer(cols = CohortQ1:ReplacedQ1,
                               names_to = 'Source',
                               values_to = 'Survival')

  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Survival, color = Source)) +
    geom_smooth(se = FALSE, linetype = 'solid', size=0.4)

  # return.object <- list(survivals, plot)
  # 
  # return(return.object)
  return(p)
}

survivalCurveReplaced(cohort, replaced1)

```
```{r "quint4"}

survivalCurveReplaced <- function(cohort, replaced) {

  # Age and survival for each quintile
  cohort$Age <- (cohort$year - 2012) + 50
  cohort$q1 <- cohort$m_endpop_wq1 / cohort$m_endpop_wq1[1]
  cohort$q4 <- cohort$m_endpop_wq4 / cohort$m_endpop_wq4[1]
  replaced$Age <- (replaced$year - 2012) + 50
  replaced$q1 <- replaced$m_endpop_wq1 / replaced$m_endpop_wq1[1]

  # Cut dataframe down to just the age and survivals
  #survivals <- cohort %>% select(Age, q1, q2, q3, q4, q5)
  survivals <- data.frame(cohort$Age, cohort$q1, cohort$q4, replaced$q1)

  survivals <- survivals %>%
                  dplyr::rename('Age' = cohort.Age, 
                                'CohortQ1' = cohort.q1, 
                                'CohortQ3' = cohort.q4, 
                                'ReplacedQ1' = replaced.q1) %>%
                  pivot_longer(cols = CohortQ1:ReplacedQ1,
                               names_to = 'Source',
                               values_to = 'Survival')

  p <- ggplot(data = survivals, mapping = aes(x = Age, y = Survival, color = Source)) +
    geom_smooth(se = FALSE, linetype = 'solid', size=0.4)

  # return.object <- list(survivals, plot)
  # 
  # return(return.object)
  return(p)
}

survivalCurveReplaced(cohort, replaced2)

### Prevalences
```{r}

subgroupPrevReplaced <- function(cohort, replaced, var, subgroups) {

  # Age
  cohort$Age <- (cohort$year - 2012) + 50
  replaced$Age <- (replaced$year - 2012) + 50

  # Build varnames
  varnames <- c()
  ticker <- 1
  for (i in subgroups) {
    varnames[ticker] <- paste0('p_', var, '_wq', i)
    print(varnames[ticker])
    ticker <- ticker + 1
  }

  # Select vars we've just built
  cohort.prev <- cohort %>% select(Age, all_of(varnames))
  replaced.prev <- cohort %>% select(Age, all_of(varnames))

  prevs <- merge(cohort.prev, replaced.prev, by = 'Age', suffixes = c('.Cohort', '.Replaced'))

  print(colnames(prevs))

  prevs <- pivot_longer(data = prevs,
                        cols = paste0(varnames[1], '.Cohort'):paste0(varnames[length(varnames)], '.Replaced'),
                        names_to = 'Source',
                        values_to = 'Prevalence')
  print(colnames(prevs))

  p <- ggplot(data = prevs, mapping = aes(x = Age, y = Prevalence, color = Source)) +
      geom_smooth(se=FALSE, size=0.4)

subgroupPrevReplaced(cohort, replaced1, var = 'anydisease', subgroups = c(1,3))
# subgroupPrevReplaced(cohort, replaced1, var = 'disabled', subgroups = c(1,3))
# subgroupPrevReplaced(cohort, replaced1, var = 'severeCondition', subgroups = c(1,3))
# subgroupPrevReplaced(cohort, replaced1, var = 'mildCondition', subgroups = c(1,3))
# 
# subgroupPrevReplaced(cohort, replaced2, var = 'anydisease', subgroups = c(1,4))
# subgroupPrevReplaced(cohort, replaced2, var = 'disabled', subgroups = c(1,4))
# subgroupPrevReplaced(cohort, replaced2, var = 'severeCondition', subgroups = c(1,4))
# subgroupPrevReplaced(cohort, replaced2, var = 'mildCondition', subgroups = c(1,4))

#someSubgroupPrevalence(cohort, var = 'anydisease', subgroups = c(1,3))

##### Any Disease
```{r}

c <- cohort
r2 <- replaced2

c$Age <- (c$year - 2012) + 50
r2$Age <- (r2$year - 2012) + 50

c.prev <- c %>% select(Age, p_anydisease_wq1, p_anydisease_wq4) %>%
                dplyr::rename('OrigPoor' = p_anydisease_wq1, 'OrigWealth' = p_anydisease_wq4)
r2.prev <- r2 %>% select(Age, p_anydisease_wq1) %>%
                dplyr::rename('ReplacedPoor' = p_anydisease_wq1)

combined <- merge(c.prev, r2.prev, by = 'Age')

combined.pivot <- pivot_longer(data = combined,
                               cols = OrigPoor:ReplacedPoor,
                               names_to = 'Source',
                               values_to = 'Prevalence')

ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
  geom_smooth(se=FALSE, size=0.4) +
  labs(title = 'Any Disease')

##### Disabled
```{r}
c <- cohort
r2 <- replaced2

c$Age <- (c$year - 2012) + 50
r2$Age <- (r2$year - 2012) + 50

c.prev <- c %>% select(Age, p_disabled_wq1, p_disabled_wq4) %>%
                dplyr::rename('OrigPoor' = p_disabled_wq1, 'OrigWealth' = p_disabled_wq4)
r2.prev <- r2 %>% select(Age, p_disabled_wq1) %>%
                dplyr::rename('ReplacedPoor' = p_disabled_wq1)

combined <- merge(c.prev, r2.prev, by = 'Age')

combined.pivot <- pivot_longer(data = combined,
                               cols = OrigPoor:ReplacedPoor,
                               names_to = 'Source',
                               values_to = 'Prevalence')

ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
  #geom_smooth(se=FALSE, size=0.4) +
  geom_line() +
  labs(title = 'Disabled')
```

##### Severe Condition
```{r}
c <- cohort
r2 <- replaced2

c$Age <- (c$year - 2012) + 50
r2$Age <- (r2$year - 2012) + 50

c.prev <- c %>% select(Age, p_severeCondition_wq1, p_severeCondition_wq4) %>%
                dplyr::rename('OrigPoor' = p_severeCondition_wq1, 'OrigWealth' = p_severeCondition_wq4)
r2.prev <- r2 %>% select(Age, p_severeCondition_wq1) %>%
                dplyr::rename('ReplacedPoor' = p_severeCondition_wq1)

combined <- merge(c.prev, r2.prev, by = 'Age')

combined.pivot <- pivot_longer(data = combined,
                               cols = OrigPoor:ReplacedPoor,
                               names_to = 'Source',
                               values_to = 'Prevalence')

ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
  geom_smooth(se=FALSE, size=0.4) +
  labs(title = 'Severe Condition')
```

##### Mild Condition
```{r}
c <- cohort
r2 <- replaced2

c$Age <- (c$year - 2012) + 50
r2$Age <- (r2$year - 2012) + 50

c.prev <- c %>% select(Age, p_mildCondition_wq1, p_mildCondition_wq4) %>%
                dplyr::rename('OrigPoor' = p_mildCondition_wq1, 'OrigWealth' = p_mildCondition_wq4)
r2.prev <- r2 %>% select(Age, p_mildCondition_wq1) %>%
                dplyr::rename('ReplacedPoor' = p_mildCondition_wq1)

combined <- merge(c.prev, r2.prev, by = 'Age')

combined.pivot <- pivot_longer(data = combined,
                               cols = OrigPoor:ReplacedPoor,
                               names_to = 'Source',
                               values_to = 'Prevalence')

ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
  geom_smooth(se=FALSE, size=0.4) +
  labs(title = 'Mild Condition')
```

##### cancre
```{r}
c <- cohort
r2 <- replaced2

c$Age <- (c$year - 2012) + 50
r2$Age <- (r2$year - 2012) + 50

c.prev <- c %>% select(Age, p_cancre_wq1, p_cancre_wq4) %>%
                dplyr::rename('OrigPoor' = p_cancre_wq1, 'OrigWealth' = p_cancre_wq4)
r2.prev <- r2 %>% select(Age, p_cancre_wq1) %>%
                dplyr::rename('ReplacedPoor' = p_cancre_wq1)

combined <- merge(c.prev, r2.prev, by = 'Age')

combined.pivot <- pivot_longer(data = combined,
                               cols = OrigPoor:ReplacedPoor,
                               names_to = 'Source',
                               values_to = 'Prevalence')

ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
  geom_smooth(se=FALSE, size=0.4)

c <- cohort
r2 <- replaced2

c$Age <- (c$year - 2012) + 50
r2$Age <- (r2$year - 2012) + 50

c.prev <- c %>% select(Age, p_hearte_wq1, p_hearte_wq4) %>%
                dplyr::rename('OrigPoor' = p_hearte_wq1, 'OrigWealth' = p_hearte_wq4)
r2.prev <- r2 %>% select(Age, p_hearte_wq1) %>%
                dplyr::rename('ReplacedPoor' = p_hearte_wq1)

combined <- merge(c.prev, r2.prev, by = 'Age')

combined.pivot <- pivot_longer(data = combined,
                               cols = OrigPoor:ReplacedPoor,
                               names_to = 'Source',
                               values_to = 'Prevalence')

ggplot(data = combined.pivot, mapping = aes(x = Age, y = Prevalence, color = Source)) +
  geom_smooth(se=FALSE, size=0.4)
```
